Algorithmic and Economic Aspects of the Internet

### 2012-12-05 report/paper.tex
Beat some things.
Signed-off-by: Sam Phippen <samphippen@googlemail.com>

Luke can you make this connect better

\subsection{AA v Shaver} One very noticable trend in our analysis is when AA
is pitched against the Shaver algorithm. In all of our experiments where trades
occurred, Shaver made more profit than AA. In this section, we explore the
reasons why Shaver so completely outclasses our implemetation of AA.

Shaver is designed to make "close shave" bids and asks by slightly out-bidding
the best bid on the market. This is similar to how AA's aggressiveness is
supposed to operate; when a bid is placed that out-bids the bot, the price is
adjusted so that the bot out-bids the previous bid in the following round
(unless such a bid would exceed the bot's limit price). This aggressiveness
behaviour increases the chance of a trade and thus increases profit.

We show 3 cases here:
\begin{enumerate}
\item 1 AA bot v. 1 Shaver bot
\item 1 AA bot v. 10 Shaver bots
\item 10 AA bots v. 10 Shaver bots
\end{enumerate}

\subsubsection{1 v. 1} No trading occurs in this experiment. In figures 2 and 3,
you can see the shouts made by the bots. The Shaver buyer and seller traders
tend toward the equilibrium price over time but don't manage to reach it before
the market session ends. The AA bot makes very few shouts and is always out-bid
/ out-asked by Shaver.

@@

\subsubsection{1 v. Many} With more Shaver bots, trading starts at around round
90. In figures 4 and 5, we see that the Shaver bots tend towards the
equilibrium as before but at a much faster rate than when there was only one
bot; this is likely due to there being 10 Shaver bots which are all trying to
out-bid each other. When the bids and asks converge, the Shaver bots invariably
trade with themselves. Our AA implementation is excluded from trading as it
can't adjust its aggressiveness quickly enough. Not all Shaver bots are engaged
in making actual trades at all times, the horizontal lines of bids / asks on the
graphs indicate that a Shaver bot is placing shouts at its limit price.

@@

\subsubsection{Many v Many} Again, trading begins at around round 90. In figures
6 and 7, we see the same pattern as we do in the previous experiment except that
AA is making more bids / asks. On this occasion we see that AA actually manages
to make a few profitable trades between rounds 90 and 140 before being swamped
by the simpler and more effective aggressiveness function of the correctly
implemented Shaver bot. These few trades are more by chance than by design; the
dip in bid price would have intersected with the poorly adapting AA bot allowing
AA to place bids that would be accepted by the seller Shaver traders.

@@

\subsection{1 v 1 tests} Given these initially dismal results, we decided that
a good place to start a detailed analysis would be a 1v1 test with the Giveaway
trader, which is clearly the worst of all the traders within the BSE system.
300 1v1 tests were performed, we used the same schedule as the many versus many
test, with a longer total trading time of 1800. Results where both bots made
zero profit were eliminated, this brought the number of tests included for
analysis down from 300 to 242. Of these tests 103 were won by Giveaway and 139
were won by our agent. The distributions of profit for each agent, as well as
the distribution of differences in profit can be seen in figures 8 and 9.

\begin{figure}[h!] \includegraphics[width=80mm]{giveaway_1_v_1.png}
\caption{Histogram of profit distributions over 300 similar trading runs for
the two bot types used in the 1v1 test. Note axis are different for the two
bots} \end{figure}

@@

From these analysis we basically conclude that whilst our bot trades more
efficiently than giveaway it does not trade more efficiently than ZIC in both
stable and extremely unstable market conditions. Whilst this is disappointing
we find an obvious reason for this result: our implementation, which attempts
to follow the maths in the original Vytelingum\cite{Vytellingum:AA} paper,
there are flaws in the implementation which cause it not to trade anywhere near
as efficiently as expected, causing it to essentially fail to trade.

Our implementation also seems to issue significantly fewer trades than would be
expected. Looking at the implementations of the other bots provided by D.
Cliff, it is not hard to see why. The implementations provided will always
return a trade when the bot has an order to fulfil (with the exception of the
Sniper bot which will make a shout on every round after a give threshold)
meaning these implementations have many more shouts on which to make profit.

This is contrary to our AA implementation which will not return a trade in the
case where the best bid in the market is greater than (for a buyer), or where
the best ask is less than (for a seller), the bot's limit price. This results
in our AA implementation not sharing it's information with the market which
would prevent other bots making offers that are more favourable to our AA bot.

\section{ZIC versus ZIP in the square wave market}

@@

In figure n we show the trades issued by each agent in one of our sample runs
with time within the trading session, along with the agent's limit price when
the trade was created.

--luke can you fill this out explaining why this happened

>>>>

\subsection{AA v Shaver} In the previous section, our AA implementation is
pitched against the Shaver algorithm. In these experiments, Shaver made more
profit than AA. In this section, we explore the reasons why Shaver so
completely outclasses our implementation of AA, causing it to make a loss in a
number of experiments.

Shaver is designed to make "close shave" bids and asks by slightly out-bidding
the best bid on the market. This is similar to how AA's aggressiveness is
supposed to operate; when a bid is placed that out-bids the agent, the price is
adjusted so that the agent out-bids the previous bid in the following round
(unless such a bid would exceed the agent's limit price). This aggressiveness
behaviour increases the chance of a trade and thus increases profit.

We show 3 cases here:
\begin{enumerate}
\item 1 AA agent v. 1 Shaver agent
\item 1 AA agent v. 10 Shaver agents
\item 10 AA agents v. 10 Shaver agents
\end{enumerate}

\subsubsection{1 v. 1} No trading occurs in this experiment. In figures 2 and 3,
you can see the shouts made by the agents. The Shaver buyer and seller traders
tend toward the equilibrium price over time but don't manage to reach it before
the market session ends. The AA agent makes very few shouts and is always out-bid
/ out-asked by Shaver.

@@

\subsubsection{1 v. Many} With more Shaver agents, trading starts at around round
90. In figures 4 and 5, we see that the Shaver agents tend towards the
equilibrium as before but at a much faster rate than when there was only one
agent; this is likely due to there being 10 Shaver agents which are all trying to
out-bid each other. When the bids and asks converge, the Shaver agents invariably
trade with themselves. Our AA implementation is excluded from trading as it
can't adjust its aggressiveness quickly enough. Not all Shaver agents are engaged
in making actual trades at all times, the horizontal lines of bids / asks on the
graphs indicate that a Shaver agents is placing shouts at its limit price.

@@

\subsubsection{Many v Many} Again, trading begins at around round 90. In figures
6 and 7, we see the same pattern as we do in the previous experiment except that
AA is making more bids / asks. On this occasion we see that AA actually manages
to make a few profitable trades between rounds 90 and 140 before being swamped
by the simpler and more effective aggressiveness function of the correctly
implemented Shaver agents. These few trades are more by chance than by design; the
dip in bid price would have intersected with the poorly adapting AA agents allowing
AA to place bids that would be accepted by the seller Shaver traders.

--conclusion paragraph here plz luke

@@

\subsection{1 v 1 tests} Given these initially dismal results, we decided that
a good place to start a detailed analysis would be a 1v1 test with the Giveaway
trader, which is clearly the worst of all the traders within the BSE system.
300 1v1 tests were performed, we used the same schedule as the many versus many
test, with a longer total trading time of 1800. Results where both agents made
zero profit were eliminated, this brought the number of tests included for
analysis down from 300 to 242. Of these tests 103 were won by Giveaway and 139
were won by our agent. The distributions of profit for each agent, as well as
the distribution of differences in profit can be seen in figures 8 and 9.

\begin{figure}[h!] \includegraphics[width=80mm]{giveaway_1_v_1.png}
\caption{Histogram of profit distributions over 300 similar trading runs for
the two agents types used in the 1v1 test. Note axis are different for the two
agents} \end{figure}

@@

From these analysis we basically conclude that whilst our agents trades more
efficiently than giveaway it does not trade more efficiently than ZIC in both
stable and extremely unstable market conditions. Whilst this is disappointing
we find an obvious reason for this result: our implementation, which attempts
to follow the maths in the original Vytelingum\cite{Vytellingum:AA} paper,
there are flaws in the implementation which cause it not to trade anywhere near
as efficiently as expected, causing it to essentially fail to trade.

Our implementation also seems to issue significantly fewer trades than would be
expected. Looking at the implementations of the other agents provided by D.
Cliff, it is not hard to see why. The implementations provided will always
return a trade when the agent has an order to fulfil (with the exception of the
Sniper agent which will make a shout on every round after a give threshold)
meaning these implementations have many more shouts on which to make profit.

This is contrary to our AA implementation which will not return a trade in the
case where the best bid in the market is greater than (for a buyer), or where
the best ask is less than (for a seller), the agent's limit price. This results
in our AA implementation not sharing it's information with the market which
would prevent other agents making offers that are more favourable to our AA agent.
\newpage
\section{ZIC versus ZIP in the square wave market}

@@

In figure 13 we show the trades issued by each agent in one of our sample runs
with time within the trading session, along with the agent's limit price when
the trade was created.

\begin{figure} \includegraphics[width=80mm]{omg_zip_zic.png}
\caption{Example of bids and asks issued by ZIC and ZIP in a market with the
conditions of this experiment}
\end{figure}

Figure 13 clearly shows why ZIC is able to outperform ZIP in this market. As the
market price rapidly shifts ZIC simply redistributes it's prices randomly
throughout the possible range of prices, ZIP's internal profit margin changes
in response to what everyone else is doing. However as the market is peturbing
randomly, and the ZIC agents are trading randomly, this estimate is likely to
be wildly thrown off. The reason ZIC can trade is that it's offers will always
eventually "meet in the middle", regardless of the, massive, random variation.
